---
title: "Assignment 2 - simple vers"
output: html_document
date: "2025-03-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install Packages 

```{r}
pacman::p_load(tidyverse, here, posterior, cmdstanr, brms, tidybayes, rstan, ggplot2, dplyr, patchwork)
set.seed(100)
```

# FUNCTIONS 

##  Agent Functions 
```{r}
# probabilistic win-stay lose-shift model (with noise)
WSLS_asym_noise <- function(prev_choice, feedback, WS_weight, LS_weight, noise){
  if(feedback == 1){  # win
    if(rbinom(1, 1, WS_weight) == 1){
      choice <- prev_choice
    } else{
      choice <- 1 - prev_choice
    }
  }
  else if (feedback == 0){  # loss
    if(rbinom(1, 1, LS_weight) == 1){
      choice <- 1 - prev_choice
    } else{
      choice <- prev_choice
    }  
  }
  # introduce noise
  if(rbinom(1, 1, noise) == 1){
    choice <- rbinom(1, 1, .5)
  }
  return(choice)
}

# random agent 
random_agent <- function(rate){
  choice <- rbinom(1, 1, rate) 
  return(choice)
} 
```

## Simulation Function
The simulation function simulates a game of MPG between a random hider and a asymmetric probabilistic WSLS guesser
```{r}
sim_MPG <- function(n, # number of trials 
                    random_rate, # rate for random hider
                    WS_weight, # rate for staying when winning for WSLS
                    LS_weight, # rate for shifting when losing for WSLS
                    noise_wsls){  # noise in WSLS
  
  random_hider <- list(choice = rep(NA, n)) #empty list of choices made by hider
  for (i in 1:n){
    random_hider$choice[i] <- random_agent(random_rate)
  }
  wsls_guesser <- list(choice = rep(NA, n), #empty list of choices made by guesser
                      feedback = rep(NA, n)) #empty list for feedback for guesser

  for(i in 1:n){
    if(i == 1){
      wsls_guesser$choice[i] <- random_agent(0.5) #random first choice 
    }
    if(i>1){
      if(wsls_guesser$choice[i-1] == random_hider$choice[i-1]){
        wsls_guesser$feedback[i-1] <- 1
      }
      else{
        wsls_guesser$feedback[i-1] <- 0
      }
      wsls_guesser$choice[i] <- WSLS_asym_noise(wsls_guesser$choice[i-1], 
                                                wsls_guesser$feedback[i-1], 
                                                WS_weight, LS_weight, noise_wsls)
    }
    if(i == n){
      if(wsls_guesser$choice[n] == random_hider$choice[n]){
        wsls_guesser$feedback[n] <- 1
      }
      else{
        wsls_guesser$feedback[n] <- 0
      }
    }
  }
  #combine data for model into list
  data <- list(n = n, choice = wsls_guesser$choice, feedback =  wsls_guesser$feedback)
  return(data)
}
```

```{r}
MPD <- function(x) {
  density(x)$x[which(density(x)$y==max(density(x)$y))]
}
```


# SIMULATE DATA 
```{r}
n <-  120
random_rate <- 0.5
WS_weight <- 0.9 
LS_weight <- 0.6 
noise <- 0
data <- sim_MPG(n, random_rate, WS_weight, LS_weight, noise)
```


# STAN code 
```{r}
stan_model <- "
data{
  int <lower=1> n; // number of trials (integers above 1)
  array[n] int <lower=0, upper=1> choice; // choice made by the guesser (length n and ints of 0 (left) or 1 (right))
  array[n] int <lower=0, upper=1> feedback; // feedback for each guess (length n and ints of 0 (loss) or 1 (win))
}
parameters{
  real theta_WS; // the bias for staying when winning on log odds scale
  real theta_LS; // the bias for shifting when losing on log odds scale
}

transformed parameters{
  vector[n] win_choice; // storing the choice when winning (0 = lost on last trial)
  vector[n] lose_choice;// storing the choice when losing  (0 = won on last trial)
  vector[n] theta;      // storing theta for each trial (probability of choosing right hand on log odds scale)
  
  theta[1] = logit(0.5);// theta for the first trial is 0.5 as there is no feedback from previous trial to use
  
  for (trial in 2:n){   // loop through each trial and define whether theta_WS or theta_LS should be used as theta
  
   if(feedback[trial-1] == 1){      // if the guesser won on the previous trial
     lose_choice[trial] = 0;        // discount theta_LS parameter when calculating theta
     
     if(choice[trial-1] == 1){      // if the guesser picked right (1) on the previous trial 
       win_choice[trial] = 1;       // use theta_WS parameter when calculating theta
     }
     else if (choice[trial-1] == 0){// if the guesser picked left (0)
       win_choice[trial] = -1;      // use theta_WS parameter when calculating theta
     }
   }
   if(feedback[trial-1] == 0){      // if the guesser lost on the previous trial  
     win_choice[trial] = 0;         // discount theta_wS parameter when calculating theta
     
     if(choice[trial-1] == 1){      // if the guesser picked right (1) on the previous trial 
       lose_choice[trial] = -1;     // use theta_LS when calculating theta
     }
     else if (choice[trial-1] == 0){// if the guesser picked left (0) on the previous trial 
       lose_choice[trial] = 1;      // use theta_LS when calculating theta 
     }
   }
   // for each trial theta is defined depending on the outcome of the previous trial (loss or win):
   theta[trial] = theta_WS*win_choice[trial] + theta_LS*lose_choice[trial]; 
  }
 
}

model{
  target += normal_lpdf(theta_WS | 0, 1); // prior for theta_WS 
  target += normal_lpdf(theta_LS | 0, 1); // prior for theta_LS
  
  // liklihood model, for each trial the choice is given the theta (as calculated in the transformed parameters chunk)
  for (trial in 1:n){
    target += bernoulli_logit_lpmf(choice[trial] | theta[trial]); 
  }
}
generated quantities{
  // Prior and Posterior theta_WS and theta_LS parameters on probability scales
  real<lower=0, upper=1> theta_WS_prior = inv_logit(normal_rng(0,1));
  real<lower=0, upper=1> theta_LS_prior = inv_logit(normal_rng(0,1));  
  real<lower=0, upper=1> theta_WS_posterior = inv_logit(theta_WS);  
  real<lower=0, upper=1> theta_LS_posterior = inv_logit(theta_LS);  
  
  // Prior predictions: // for each trial estimate the choice based on prior theta 
  vector[n] prior_preds;                   // array for prior predictions
  vector[n] prior_theta_l;                 // array for theta on log odds scale based on priors
  real theta_WS_prior_l = normal_rng(0,1); // theta_WS prior on log-odds scale
  real theta_LS_prior_l = normal_rng(0,1); // theta_LS prior on log_odds scale

  prior_theta_l[1] = logit(0.5); // prior theta for the first trial is 0.5 (random)

  prior_theta_l[2:n] = theta_WS_prior_l*win_choice[2:n] + theta_LS_prior_l*lose_choice[2:n]; 

  for (trial in 1:n){       
    prior_preds[trial] = binomial_rng(1, inv_logit(prior_theta_l[trial])); 
  }
 
  // Posterior predictions: for each trial estimate the choice based on theta
  array[n] real posterior_preds; // array for posterior predictions 
  for (trial in 1:n){       
    posterior_preds[trial] = binomial_rng(1, inv_logit(theta[trial])); 
  }
  
}
"
write_stan_file(
  stan_model, dir = "stan/", basename = "WSLSasym.stan")
```

# COMPILE MODEL 
```{r}
file <- file.path("stan/WSLSasym.stan")
mod <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE))

samples <- mod$sample(
  data = data,
  seed = 120,
  refresh = 500,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  max_treedepth = 20,
  adapt_delta = 0.99
)

samples$summary()
draws_df <- as_draws_df(samples$draws())
```

# PLOT MODEL QUALITY CHECKS
## PRIOR PREDICTIVE CHEKCKS & POSTERIOR PREDICTIVE CHECKS
```{r}
# Posterior Predictions
posterior_predictions <- draws_df %>%
  select(starts_with("posterior_preds[")) %>% 
  pivot_longer(everything(), 
              names_to = "trial",
              values_to = "prediction") %>%
  mutate(trial = as.numeric(str_extract(trial, "\\d+")))

posterior_summary <- posterior_predictions %>%
  group_by(trial) %>%
  summarise(
    mean = mean(prediction),
    lower = quantile(prediction, 0.025),
    upper = quantile(prediction, 0.975))

# Prior Predictions
prior_predictions <- draws_df %>%
  select(starts_with("prior_preds[")) %>%
  pivot_longer(everything(), 
              names_to = "trial",
              values_to = "prediction") %>%
  mutate(trial = as.numeric(str_extract(trial, "\\d+")))

prior_summary <- prior_predictions %>%
  group_by(trial) %>%
  summarise(
    mean = mean(prediction),
    lower = quantile(prediction, 0.025),
    upper = quantile(prediction, 0.975))

# Plot: Prior Predictive Check
p1 <- ggplot() +
  # Interval
  geom_ribbon(data = prior_summary, aes(x = trial, ymin = lower, ymax = upper),
              alpha = 0.2, fill = "red") +
  # Mean Prior Prediction 
  geom_line(data = prior_summary, aes(x = trial, y = mean), color = "red") +
  # True choices
  geom_point(data = tibble(trial = 1:data$n, choice = data$choice),
             aes(x = trial, y = choice), alpha = 0.5) +
  labs(title = "Prior Predictive Check", x = "Trial", y = "Choice (0/1)") +
  theme_minimal()

# Plot: Posterior Predictive Check
p2 <- ggplot() +
  # Interval
  geom_ribbon(data = posterior_summary, aes(x = trial, ymin = lower, ymax = upper),
              alpha = 0.2, fill = "blue") +
# Mean Posterior Prediction
  geom_line(data = posterior_summary, aes(x = trial, y = mean), color = "blue") +
  # True Choices
  geom_point(data = tibble(trial = 1:data$n, choice = data$choice),
             aes(x = trial, y = choice), alpha = 0.5) +
  labs(title = "Posterior Predictive Check", x = "Trial", y = "Choice (0/1)") +
  theme_minimal()
p1+p2

ggsave("figs/posterior_&_prior_predictive_check.png", width = 14, height = 7)
```


## PRIOR-POSTERIOR UPDATE CHECKS 
```{r}
ggplot(draws_df) +
  geom_density(aes(theta_WS_posterior), fill = "blue", alpha = 0.3) +
  geom_density(aes(theta_WS_prior), fill = "red", alpha = 0.3) +
  geom_vline(xintercept = WS_weight, linetype = "dashed", color = "black", size = 1) +
  ggtitle("Prior Posterior Update Checks: THETA WIN-STAY")+
  xlab("Rate") +
  ylab("Estimate Densities") +
  theme_classic()
ggsave("figs/prior_post_update_thetaWS.png")

ggplot(draws_df) +
  geom_density(aes(theta_LS_posterior), fill = "blue", alpha = 0.3) +
  geom_density(aes(theta_LS_prior), fill = "red", alpha = 0.3) +
  geom_vline(xintercept = LS_weight, linetype = "dashed", color = "black", size = 1) +
  ggtitle("Prior Posterior Update Checks: THETA LOSE-SHIFT")+
  xlab("Rate") +
  ylab("Estimate Densities") +
  theme_classic()
ggsave("figs/prior_post_update_thetaLS.png")
```


## MARKOV CHAINS
```{r}
ggplot(draws_df, aes(.iteration, theta_WS_posterior, group = .chain, color = .chain)) +
  geom_line() +
  ggtitle("THETA WIN-STAY")+
  theme_classic()
ggsave("figs/markov_chains_WS.png")

ggplot(draws_df, aes(.iteration, theta_LS_posterior, group = .chain, color = .chain)) +
  geom_line() +
  ggtitle("THETA LOSE-SHIFT")+
  theme_classic()
ggsave("figs/markov_chains_LS.png")
```

# PARAMETER RECOVERY

## Simulate Data with different theta values and different number of trials
theta_WS are bounded between 0.5 (so it mostly stays) and 0.99 (no extreme values)
theta_LS are bounded between 0.5 (so it mostly shift) and 0.99 (no extreme values)
number of trials: 30, 60, 90, 120, 150

for each run the two theta parameters are generated and simulated for each of the 5 possible trial numbers

```{r}
nruns <- 100
true_theta <- list(true_theta_WS = rep(NA, nruns), true_theta_LS = rep(NA, nruns))
ntrials <- c(30, 60, 90, 120, 150)
data_list <- vector("list", nruns)

for(i in 1:nruns){
  true_theta$true_theta_WS[i] <- runif(1, min = 0.5, max = 0.99)
  true_theta$true_theta_LS[i] <- runif(1, min = 0.5, max = 0.99)
  
  data_sublist <- vector("list", length(ntrials))

  for(j in 1:length(ntrials)){
    data_sublist[[j]] <- sim_MPG(n = ntrials[j], 
                                random_rate = 0.5, 
                                WS_weight = true_theta$true_theta_WS[i] , 
                                LS_weight = true_theta$true_theta_LS[i] , 
                                noise_wsls = 0)
  }
  data_list[[i]] <- data_sublist
}
```

## Fit the model to each run of the simulated data
```{r}
recovery_df <- NULL

for(i in 1:nruns){
  
  data_sub <- data_list[[i]] #subset with five sims in 
  
  for(j in 1:length(ntrials)){
    data <- data_sub[[j]]
    
    samples <- mod$sample(
      data = data,
      seed = 123,
      refresh = 500,
      chains = 2,
      parallel_chains = 2,
      threads_per_chain = 2,
      iter_warmup = 1000,
      iter_sampling = 2000,
      max_treedepth = 20,
      adapt_delta = 0.99
    )
    
    draws_df <- as_draws_df(samples$draws()) 
    
    temp <- tibble(# true theta values 
                   true_theta_WS = true_theta$true_theta_WS[i],
                   true_theta_LS = true_theta$true_theta_LS[i],
                   # estimated theta values
                   est_theta_WS = draws_df$theta_WS_posterior, 
                   est_theta_LS = draws_df$theta_LS_posterior, 
                   # maximum posterior densities for theta values
                   est_theta_WS_mpd = MPD(draws_df$theta_WS_posterior), 
                   est_theta_LS_mpd = MPD(draws_df$theta_LS_posterior),
                   #number of trials 
                   n = data$n)
    
    if(exists("recovery_df")){
      recovery_df <- rbind(recovery_df, temp)
    } 
    else{
      recovery_df <- temp
    }
  }
}
save(recovery_df, file = "RData/recovery_estimates.RData")
```


## Load saved recovery data
```{r}
load("RData/recovery_estimates.RData")
```

## Plot Recovery: Full range of estimated values(n = 120)
```{r}
recovery_df %>% 
  filter(n == 120) %>% 
  ggplot(aes(true_theta_WS, est_theta_WS)) +
    geom_point(alpha = 0.1) +
    geom_smooth() +
    ggtitle("Parameter Recovery: THETA WIN-STAY (n=120)")+
    theme_minimal()
ggsave("figs/recovery_theta_WS.png")

recovery_df %>% 
  filter(n == 120) %>% 
  ggplot(aes(true_theta_LS, est_theta_LS)) +
    geom_point(alpha = 0.1) +
    geom_smooth() +
    ggtitle("Parameter Recovery: THETA LOSE-SHIFT (n=120)")+
    theme_minimal()
ggsave("figs/recovery_theta_LS.png")

```


## Plot Recovery: Maximum Posterior Density
```{r}
recovery_df %>% 
  filter(n == 120) %>% 
  ggplot(aes(true_theta_WS, est_theta_WS_mpd)) +
    geom_point(alpha = 0.1) +
    #xlim(0.5,1) + ylim(0.5,1)+
    geom_abline(aes(slope = 1, intercept = 0, colour = "red"), show.legend = F)+
    ggtitle("Parameter Recovery (MPD): THETA WIN-STAY (n=120)")+
    theme_minimal()
ggsave("figs/recovery_theta_WS_mpd.png")

recovery_df %>% 
  filter(n == 120) %>%  
  ggplot(aes(true_theta_LS, est_theta_LS_mpd)) +
    geom_point(alpha = 0.1) +
    #xlim(0,1) + ylim(0,1)+
    geom_abline(aes(slope = 1, intercept = 0, colour = "red"), show.legend = F)+
    ggtitle("Parameter Recovery (MPD): THETA LOSE-SHIFT (n=120)")+
    theme_minimal()
ggsave("figs/recovery_theta_LS_mpd.png")

```
## Plot: Recovery per number of trials

```{r}
 
ggplot(recovery_df, aes(true_theta_LS, est_theta_LS_mpd)) +
  geom_point(alpha = 0.1) +
  geom_abline(aes(slope = 1, intercept = 0, colour = "red"), show.legend = F)+
  ggtitle("Parameter Recovery: THETA LOSE-SHIFT")+
  facet_wrap(~n, labeller = label_both)+
  theme_minimal()
ggsave("figs/recovery_theta_LS_trials.png")


ggplot(recovery_df, aes(true_theta_WS, est_theta_WS_mpd)) +
  geom_point(alpha = 0.1) +
  #xlim(0,1) + ylim(0,1)+
  geom_abline(aes(slope = 1, intercept = 0, colour = "red"), show.legend = F)+
  ggtitle("Parameter Recovery: THETA WIN-STAY")+
  facet_wrap(~n, labeller = label_both)+
  theme_minimal()
ggsave("figs/recovery_theta_WS_trials.png")

```


