---
title: "Assignment 1"
course: "Advanded Cognitive Modeling"
output: html_document
date: "2025-02-11"
authors: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
pacman::p_load(tidyverse, tidyr, dplyr, ggplot2)
```

# Strategy functions 

Implemented strategies are formalised in the .R script 'functions.R'. 
Strategy 1 is an asymmetric and probabilistic version of the 'win-stay lose-shift' strategy, with noise. 
Strategy 2 is a memory model with a moving average, taking previous 7 trials into account (this also includes noise).
Additionally, a deterministic 'win-stay lose-shift' strategy (strategy 3) is formalised and used as the hider in both run 1 (against strategy 1) and run 2 (against strategy 2). 
The choices in the first trial for all runs and strategies are determined using a random strategy. 
```{r}
source("functions.R")
```

# Running the Matching Pennies Game 
Two runs are setup with some shared parameters (and other run specific). 
Each run have 100 agents (or pairs of agents) and each pair 'play' for a 120 trials. The noise parameter for all models are set to the same, 0.1. 
## Parameters 
```{r}
noise <- 0.1
n_trials <- 120
n_agents <- 100
```

## Run number 1: Strategy 1 versus Strategy 3

### Simulation
```{r}
WS_weight <- 0.9 #probability of staying when winning str1
LS_weight <- 0.6 #probability of shifting when losing str1

run1 <- array(NA, c(n_trials, n_agents))

for (a in 1:n_agents){
  
  guesser_asym_wsls <- rep(NA, n_trials)
  hider_wsls <-  rep(NA, n_trials)

  #first trial is based on nothing, i.e. random guess/hide
  guesser_asym_wsls[1] <- random_agent(0.5) 
  hider_wsls[1] <- random_agent(0.5) 

  for (i in 2:n_trials){
    if(guesser_asym_wsls[i-1] == hider_wsls[i-1]){
      feedback <- 1
    }
    else{
      feedback <- 0
    }
    guesser_asym_wsls[i] <- WSLS_asym_noise(guesser_asym_wsls[i-1], feedback, 
                                            WS_weight, LS_weight, noise)
    hider_wsls[i] <- WSLS_noise(hider_wsls[i-1], 1-feedback, noise)
  }
  
  run1[,a] <- as.numeric(guesser_asym_wsls == hider_wsls)
}
```

### Data frames
Wrangling the data from the simulation (feedback) for plotting
```{r}
#using only the last simulation as an example: 
df1_1 <- tibble(guesser_asym_wsls, hider_wsls, trial = seq(n_trials), 
             feedback = as.numeric(guesser_asym_wsls == hider_wsls))
df1_2 <- df1_1 %>% 
  mutate(cum_guesser = cumsum(feedback) / seq_along(feedback),
         cum_hider = cumsum(1 - feedback) / seq_along(feedback))

# Using all 100 agents: 
# Convert run1 (from "wide" to "long" format)
df1_agents <- as.data.frame(run1)
df1_agents$trial <- seq_len(n_trials)
df1_long <- df1_agents %>%
  pivot_longer(cols = -trial, names_to = "agent", values_to = "feedback") %>%
  group_by(agent) %>%
  mutate(cum_guesser = cumsum(feedback) / seq_along(feedback),
         cum_hider = cumsum(1 - feedback) / seq_along(feedback))  # the inverse

# Compute the average cumulative performance for both roles
df1_avg <- df1_long %>%
  group_by(trial) %>%
  summarise(avg_guesser = mean(cum_guesser), avg_hider = mean(cum_hider))


#Legend colors for plots
colors <- c("Guesser: prob-WSLS" = "red", "Average Guesser" = "red", 
            "Hider: determ-WSLS" = "blue", "Average Hider" = "blue")

```

### Plot of the last simulated agent pair
```{r}
ggplot(df1_1)+ 
  geom_line(aes(trial, guesser_asym_wsls, color = "Guesser: prob-WSLS",)) +
  geom_line(aes(trial, hider_wsls, color ="Hider: determ-WSLS"))+
  labs(title = "Behaviour Plot: Probabilistic versus Deterministc WSLS",
    x = "Trial", y = "Choice (0/1)", color = "Color legend")+
  scale_color_manual(values = colors) + 
  theme_minimal()
ggsave("figs/run1_oneagent_beh.png", width = 10, height = 7)

ggplot(df1_2)+ 
  geom_line(aes(trial, cum_guesser, color = "Guesser: prob-WSLS",)) +
  geom_line(aes(trial, cum_hider, color ="Hider: determ-WSLS"))+
  labs(title = "Performance Plot: Probabilistic versus Deterministc WSLS",
    x = "Trial", y = "Choice (0/1)", color = "Color legend")+
  scale_color_manual(values = colors) + 
  theme_minimal()
ggsave("figs/run1_oneagent_perf.png", width = 10, height = 7)
```


### Plot of 100 simulated agent pairs (and averages)
```{r}
# Plot all individual agents with low alpha and overlay the average
ggplot(df1_long) +
  # individual guesser and average guesser
  geom_line(aes(x = trial, y = cum_guesser, group = agent,  
                color = "Guesser: prob-WSLS"), alpha = 0.05) +
  geom_line(data = df1_avg, aes(x = trial, y = avg_guesser,  
                               color = "Average Guesser"), size = 1) +
  # individual hider and average hider
  geom_line(aes(x = trial, y = cum_hider, group = agent, 
                color = "Hider: determ-WSLS"), alpha = 0.05) +
  geom_line(data = df1_avg, aes(x = trial, y = avg_hider, 
                               color = "Average Hider"), size = 1) +
  scale_x_continuous(limits = c(0, 120), breaks = seq(0, 120, by = 20)) +
  labs(title = "Performance Plot: Probabilistic versus Deterministc WSLS",
       x = "Trial", y = "Cumulative Performance", color = "")+
  scale_color_manual(values = colors) + 
  theme_minimal()
ggsave("figs/run1_100agents_perf.png", width = 10, height = 7)
```



## Run number 2: Strategy 2 versus Strategy 3

### Simulation 

```{r}
run2 <- array(NA, c(n_trials, n_agents))

for (a in 1:n_agents){
  
  #empty lists for storing choices for each agent
  guesser_memory <- rep(NA, n_trials)
  hider_wsls <-  rep(NA, n_trials)
  #first trial is based on nothing, i.e. random guess/hide
  guesser_memory[1] <- random_agent(0.5) 
  hider_wsls[1] <- random_agent(0.5) 

  for (i in 2:n_trials){
    if(guesser_memory[i-1] == hider_wsls[i-1]){
      feedback <- 1
    }
    else{
      feedback <- 0
    }
    hider_choices7 <- hider_wsls %>% na.omit() %>% rev()
    hider_choices7 <- hider_choices7[1:7] %>% na.omit()#hiders choices on the previous 7 trials
    
    guesser_memory[i] <- memory_agent(hider_choices7, noise)
    hider_wsls[i] <- WSLS_noise(hider_wsls[i-1], 1-feedback, noise)
  }
  
  run2[, a] <- as.numeric(guesser_memory == hider_wsls)
}
```

### Data frames
Wrangling the data from the simulation (feedback) for plotting
```{r}
#using only the last simulation as an example: 
df2_1 <- tibble(guesser_memory, hider_wsls, trial = seq(n_trials), 
             feedback = as.numeric(guesser_memory == hider_wsls))
df2_2 <- df2_1 %>% 
  mutate(cum_guesser = cumsum(feedback) / seq_along(feedback),
         cum_hider = cumsum(1 - feedback) / seq_along(feedback))

# Using all 100 agents: 
# Convert run1 (from "wide" to "long" format)
df2_agents <- as.data.frame(run2)
df2_agents$trial <- seq_len(n_trials)
df2_long <- df2_agents %>%
  pivot_longer(cols = -trial, names_to = "agent", values_to = "feedback") %>%
  group_by(agent) %>%
  mutate(cum_guesser = cumsum(feedback) / seq_along(feedback),
         cum_hider = cumsum(1 - feedback) / seq_along(feedback))  # the inverse

# Compute the average cumulative performance for both roles
df2_avg <- df2_long %>%
  group_by(trial) %>%
  summarise(avg_guesser = mean(cum_guesser), avg_hider = mean(cum_hider))


#Legend colors for plots
colors2 <- c("Guesser: memory7" = "red", "Average Guesser" = "red", 
            "Hider: determ-WSLS" = "blue", "Average Hider" = "blue")

```

### Plot of the last simulated agent pair
```{r}
ggplot(df2_1)+ 
  geom_line(aes(trial, guesser_memory, color = "Guesser: memory7",)) +
  geom_line(aes(trial, hider_wsls, color ="Hider: determ-WSLS"))+
  labs(title = "Behaviour Plot: Memory model (7) versus Deterministc WSLS",
    x = "Trial", y = "Choice (0/1)", color = "Color legend")+
  scale_color_manual(values = colors2) + 
  theme_minimal()
ggsave("figs/run2_oneagent_beh.png", width = 10, height = 7)

ggplot(df2_2)+ 
  geom_line(aes(trial, cum_guesser, color = "Guesser: memory7",)) +
  geom_line(aes(trial, cum_hider, color ="Hider: determ-WSLS"))+
  labs(title = "Performance Plot: Memory model (7) versus Deterministc WSLS",
    x = "Trial", y = "Choice (0/1)", color = "Color legend")+
  scale_color_manual(values = colors2) + 
  theme_minimal()
ggsave("figs/run2_oneagent_perf.png", width = 10, height = 7)
```


### Plot of 100 simulated agent pairs (and averages)
```{r}

# Plot all individual agents with low alpha and overlay the average
ggplot(df2_long) +
  # individual guesser and average guesser
  geom_line(aes(x = trial, y = cum_guesser, group = agent,  
                color = "Guesser: memory7"), alpha = 0.05) +
  geom_line(data = df2_avg, aes(x = trial, y = avg_guesser,  
                               color = "Average Guesser"), size = 1) +
  # individual hider and average hider
  geom_line(aes(x = trial, y = cum_hider, group = agent, 
                color = "Hider: determ-WSLS"), alpha = 0.05) +
  geom_line(data = df2_avg, aes(x = trial, y = avg_hider, 
                               color = "Average Hider"), size = 1) +
  scale_x_continuous(limits = c(0, 120), breaks = seq(0, 120, by = 20)) +
  labs(title = "Performance Plot: Memory model (7) versus Deterministc WSLS",
       x = "Trial", y = "Cumulative Performance", color = "")+
  scale_color_manual(values = colors2) + 
  theme_minimal()
ggsave("figs/run2_100agents_perf.png", width = 10, height = 7)
```




```{r}

##fix the hider list so it doesnt include NAs 
## feedback in run1 input to WSLS_noise model has to be inversed
## I have changed the memory model 


```




