---
title: "Assignment 3"
output: html_document
date: "2025-03-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set up
```{r}
pacman::p_load(tidyverse, here, posterior, cmdstanr, brms, tidybayes, rstan, 
               ggplot2, dplyr, patchwork)

```


# FUNCTIONS
## Function for simulating first and group rating for one agent for given amount
of stimuli (trials)
```{r}
simulating_ratings <- function(n_stim, max_rating){
  
  ratings <- list(first_rating = rep(NA, n_stim), 
                  group_rating = rep(NA, n_stim))
  
  for (i in 1:n_stim){
    # simulating the first rating based on a randomly sampled probability value
    prob <- runif(1, min = 0.001, max = 0.999) 
    ratings$first_rating[i] <- rbinom(1, (max_rating), prob) #rbinom(1,(max_rating-1),prob)+1
    
    # simulating the group rating, based on the value of the first rating 
    feedback <- (sample.int((max_rating), 1) - 4) # (sample.int((max_rating - 1), 1) - 4)
    group_rating <- ratings$first_rating[i] + (feedback)
    ratings$group_rating[i] <- ifelse(group_rating < 0, 0, 
                                      ifelse(group_rating > max_rating, max_rating,
                                             group_rating))
    
    #obs: feedback is constrained to not go outside the bounds of rating scale 
    #(0-8). we should do it in another way because this way creates uneven dist
    #of feedback values (below 0 is changed to 0 and above 8 is changed to 8, 
    #creating too many values of 0 and 8). 
  }
  return(ratings)
}
```

## Beta Binomial Agent - simple model
Simulating second_rating!
IWeight parameters are included in the  function, if none weighted agent weights = 1.
```{r}
betaBinomialModel_simple <- function(alpha_prior, beta_prior, first_ratings, group_ratings, 
                                      max_rating, n_stim, weight_first, weight_group) {
  #list of values 
  ratings2 <- list(second_rating = rep(NA, n_stim), 
                   expected_rate = rep(NA, n_stim), 
                   alpha_post = rep(NA, n_stim), 
                   beta_post = rep(NA, n_stim))
  #for each stimuli: 
  for (i in 1:n_stim){
    neg_first_rating <- max_rating - first_ratings[i] #the 'points' not given 
    neg_group_rating <- max_rating - group_ratings[i] #
    
    #calculate weighted information 
    first_rating_w <- first_ratings[i] * weight_first
    group_rating_w <- group_ratings[i] * weight_group
    neg_first_rating_w <- neg_first_rating * weight_first
    neg_group_rating_w <- neg_group_rating * weight_group
  
    # calculate posterior beta and alpha values
    ratings2$alpha_post[i] <- alpha_prior + first_rating_w + group_rating_w
    ratings2$beta_post[i] <- beta_prior + neg_first_rating_w + neg_group_rating_w      
    
    # calculated second rating based on first and group rating
    ratings2$expected_rate[i] <- ratings2$alpha_post[i] /(
                                        ratings2$alpha_post[i] + ratings2$beta_post[i])  
    ratings2$second_rating[i] <- rbinom(1, 7, ratings2$expected_rate[i]) #+1
  }
  return(ratings2)
}
```

## Beta Binomial Agent - weighted model
```{r}
betaBinomialModel_weighted <- function(alpha_prior, beta_prior, first_ratings, group_ratings, 
                                        max_rating, n_stim, total_weight, weight_proportion) {
  #list of values 
  ratings2 <- list(second_rating = rep(NA, n_stim), 
                   expected_rate = rep(NA, n_stim), 
                   alpha_post = rep(NA, n_stim), 
                   beta_post = rep(NA, n_stim), 
                   total_weight = total_weight, 
                   weight_proportion = weight_proportion)
  
  weight_first <- total_weight * weight_proportion
  weight_group <- total_weight * (1 - weight_proportion)
  
  #for each stimuli: 
  for (i in 1:n_stim){
    neg_first_rating <- max_rating - first_ratings[i]
    neg_group_rating <- max_rating - group_ratings[i]
    
    #calculate weighted information 
    first_rating_w <- first_ratings[i] * weight_first
    group_rating_w <- group_ratings[i] * weight_group
    neg_first_rating_w <- neg_first_rating * weight_first
    neg_group_rating_w <- neg_group_rating * weight_group
  
    # calculate posterior beta and alpha values
    ratings2$alpha_post[i] <- alpha_prior + first_rating_w + group_rating_w
    ratings2$beta_post[i] <- beta_prior + neg_first_rating_w + neg_group_rating_w      
    
    # calculated second rating based on first and group rating
    ratings2$expected_rate[i] <- ratings2$alpha_post[i] /(
                                        ratings2$alpha_post[i] + ratings2$beta_post[i])  
    ratings2$second_rating[i] <- rbinom(1, 7, ratings2$expected_rate[i]) #+1
  }
  return(ratings2)
}

```


# SIMULATE DATA
## Simple (non weighted) simulation
```{r}
n_stim <- 153 # number of stimuli (faces)
max_rating <- 7 # maximum possible rating
n_agents <- 10 # number of agents

ratings_all <- vector("list", n_agents) #list for first and group rating for all agents 
second_ratings_all <- vector("list", n_agents) #list for second rating for all agents 

#loop through each agent
for (a in 1:n_agents){
  ratings_all[[a]] <-  simulating_ratings(n_stim, max_rating) #simulate first and group ratings
  
  second_ratings_all[[a]] <- betaBinomialModel_simple(alpha_prior = 1, beta_prior = 1,
                                               first_rating = ratings_all[[a]]$first_rating, 
                                               group_rating = ratings_all[[a]]$group_rating,
                                               max_rating, n_stim, 
                                               weight_first = 1, weight_group = 1)
}

#save data in STAN type of list 
data_simple <- list(n_stim = 153, 
                    n_agents = 10, 
                    max_rating = 7,
                    second_rating = matrix(unlist(do.call(rbind, second_ratings_all)[ 
                      ,"second_rating"]), byrow = T, ncol = n_agents),
                    first_rating = matrix(unlist(do.call(rbind, ratings_all)[ 
                      ,"first_rating"]), byrow = T, ncol = n_agents), 
                    group_rating =  matrix(unlist(do.call(rbind, ratings_all)[ 
                      ,"group_rating"]), byrow = T, ncol = n_agents))

```

## Weighted simulation
```{r}
n_stim <- 153 # number of stimuli (faces)
n_populations <- 1 # number of populations
max_rating <- 7 # maximum possible rating
n_agents <- 100 # number of agents

ratings_all <- vector("list", n_agents) #list for first and group rating for all agents 
second_ratings_all <- vector("list", n_agents) #list for second rating for all agents 

weight <- list(total_weight_mu = rep(NA, n_populations), 
               total_weight_sigma = rep(NA, n_populations), 
               weight_proportion_mu = rep(NA, n_populations),
               weight_proportion_sigma = rep(NA, n_populations))

# loop through the population
for (p in 1:n_populations){
  
  total_weight_mu <- 10
  total_weight_sigma <- 2
  
  weight_proportion_mu <- 0.8
  weight_proportion_sigma <- 0.1
  
  #loop through each agent
  for (a in 1:n_agents){
    
    total_weight <- rnorm(1, mean = total_weight_mu, sd = total_weight_sigma)
    weight_proportion <- rnorm(1, mean = weight_proportion_mu, sd = weight_proportion_sigma)
    
    ratings_all[[a]] <-  simulating_ratings(n_stim, max_rating) #simulate first and group ratings
    
    second_ratings_all[[a]] <- betaBinomialModel_weighted(alpha_prior = 1, beta_prior = 1,
                                                 first_rating = ratings_all[[a]]$first_rating, 
                                                 group_rating = ratings_all[[a]]$group_rating,
                                                 max_rating, n_stim, 
                                                 total_weight = total_weight, 
                                                 weight_proportion = weight_proportion)
  }
  weight$total_weight_mu[p] <- total_weight_mu
  weight$total_weight_sigma[p] <- total_weight_sigma
  weight$weight_proportion_mu[p] <- weight_proportion_mu
  weight$weight_proportion_sigma[p] <- weight_proportion_sigma
}

#save data in STAN type of list 
data_weighted <- list(n_stim = 153, 
                    n_agents = 100, 
                    n_populations = 1,
                    max_rating = 7,
                    second_rating = matrix(unlist(do.call(rbind, second_ratings_all)[ 
                      ,"second_rating"]), byrow = T, ncol = n_agents),
                    first_rating = matrix(unlist(do.call(rbind, ratings_all)[ 
                      ,"first_rating"]), byrow = T, ncol = n_agents), 
                    group_rating =  matrix(unlist(do.call(rbind, ratings_all)[ 
                      ,"group_rating"]), byrow = T, ncol = n_agents))
```


```{r}
hist(second_ratings_all)
```


# STAN MODEL of the simple agent model
```{r}
stan_model_simple <- "

data {
  int <lower=1> n_stim;
  int <lower=1> n_agents;
  int <lower=1> max_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> first_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> second_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> group_rating;
}

parameters{
  real alpha_prior;
  real beta_prior;
}

model {
  for (i in 1:n_agents) {
    for (j in 1:n_stim) {
      
      real alpha_post = alpha_prior + first_rating[j, i] + group_rating[j, i];
      real beta_post = beta_prior + (max_rating - first_rating[j, i]) + (max_rating - group_rating[j, i]);
      
      // model the second rating
      target += beta_binomial_lpmf(second_rating[j, i] | 7, alpha_post, beta_post);
    }
  }
}
generated quantities {

  array[n_stim, n_agents] real log_lik;
  
  array[n_stim, n_agents] int prior_pred_rating;
  array[n_stim, n_agents] int post_pred_rating;

  for (i in 1:n_agents) {
    for (j in 1:n_stim) {
      
      prior_pred_rating[j, i] = beta_binomial_rng(7, 1, 1);
      
      real alpha_post = alpha_prior + first_rating[j, i] + group_rating[j, i];
      real beta_post = beta_prior+(max_rating - first_rating[j, i]) + (max_rating - group_rating[j, i]);
      
      post_pred_rating[j, i] = beta_binomial_rng(7, alpha_post, beta_post);
      
      log_lik[j, i] = beta_binomial_lpmf(second_rating[j, i] | 7, alpha_post, beta_post);
    }
  }
}
"

write_stan_file(
  stan_model_simple, dir = "stan/", basename = "simple_agent.stan")

```

# COMPILE MODEL 
```{r}
file <- file.path("stan/simple_agent.stan")
mod <- cmdstan_model(file, cpp_options = list(stan_threads = TRUE))

samples_simple <- mod$sample(
  data = data_simple,
  seed = 120,
  refresh = 500,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  max_treedepth = 20,
  adapt_delta = 0.99
)

```

```{r}
samples_simple$summary('prior_pred_rating') # prior_pred_rating, post_pred_rating

draws_df_simple <- as_draws_df(samples_simple$draws())
draws_df_simple
```


# STAN MODEL of the weighted agent model
```{r}
stan_model_weighted <- "

data {
  int <lower=1> n_stim;
  int <lower=1> n_populations;
  int <lower=1> n_agents;
  int <lower=1> max_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> first_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> second_rating;
  array[n_stim, n_agents] int <lower=0, upper=7> group_rating;
}

parameters{
  array[n_agents] real total_weight;
  array[n_agents] real weight_proportion;
  
  real total_weight_mu;
  real total_weight_sigma;
  real <lower=0, upper=1> weight_proportion_mu;
  real weight_proportion_sigma;
}

trasnformed parameters {
  real<lower=0> weight_first = total_weight * weight_proportion;
  real<lower=0> weight_group = total_weight * (1 - weight_proportion);
}

model {
  // priors
  target += lognormal_lpdf(total_weight | )
  target += beta_lpdf(weight_proportion | )
  

  for (i in 1:n_agents) {
    for (j in 1:n_stim) {
      
      real alpha_post = alpha_prior + first_rating[j, i] + group_rating[j, i];
      real beta_post = beta_prior + (max_rating - first_rating[j, i]) + (max_rating - group_rating[j, i]);
      
      // model the second rating
      target += beta_binomial_lpmf(second_rating[j, i] | 7, alpha_post, beta_post);
    }
  }
}

generated quantities {

}
"

write_stan_file(
  stan_model_weighted, dir = "stan/", basename = "weighted_agent.stan")

```




